% ============================================================================
% Method (paper body). Written to match the implemented model in
% model/oaa_fullres.py (class OAAv2Depth, full-resolution decoder-upgrade variant).
% Requires: \usepackage{amsmath}. Notation: R=rays, N=views, M=mic tokens, C=width.
% ============================================================================
\section{Method}
\label{sec:method}

\subsection{Problem setup}
A device at a fixed position emits a short sweep and records the binaural response at
one or more known yaw orientations. Stacking the two ears of each recorded
orientation gives $N$ \emph{views}, indexed by a pose
$\pi_i=(\psi_i,\epsilon_i)$ with yaw $\psi_i$ and ear sign $\epsilon_i\!\in\!\{-1,+1\}$
($-1{=}$left, $+1{=}$right). Each view is a magnitude STFT
$\mathbf{s}_i\!\in\!\mathbb{R}^{1\times H\times W}$ ($H{\times}W{=}256{\times}512$).
The goal is a per-ray \emph{radial} depth map
$\mathbf{d}\!\in\![0,d_{\max}]^{256\times512}$ over the equirectangular (ERP) sphere.
Because the poses are known, the task is not to reconstruct stereo geometry but to
\emph{fuse pose-aligned views}; our design makes this alignment explicit.

\subsection{Per-view weight-shared encoder}
All $N$ views are encoded by the \emph{same} 2-D encoder $E$, applied with the view
axis folded onto the batch (Fig.~\ref{fig:arch}). $E$ begins with a learned strided
stem that ingests the input at full $256\times512$ resolution---so the time axis,
which carries echo delay, is seen before any pooling---and produces a coarse token
grid of size $L_h\times L_w=16\times32$ with width $C$:
\begin{equation}
\mathbf{t}_i = E(\mathbf{s}_i)\in\mathbb{R}^{L_hL_w\times C},\qquad i=1,\dots,N .
\end{equation}
Sharing $E$ across views makes the parameter count \emph{independent of $N$}: adding
orientations costs no extra weights, which is what lets the model scale from 2 to 8
channels without growth.

\subsection{Pose conditioning}
Each pose is embedded as $\mathbf{p}_i=[\sin\psi_i,\cos\psi_i,\epsilon_i]$ and injected
in two complementary ways. First, an additive pose embedding is added to the view
tokens. Second, and more importantly, an \emph{adaptive LayerNorm} (AdaLN) predicts
per-view scale/shift $(\gamma_i,\beta_i)=\mathrm{MLP}(\mathbf{p}_i)$ that modulate the
normalisation inside every view-conditioned self-attention block,
\begin{equation}
\mathrm{AdaLN}(\mathbf{x};\mathbf{p}_i)=\gamma_i\odot\mathrm{LN}(\mathbf{x})+\beta_i .
\end{equation}
AdaLN is the single architectural change that produces a consistent accuracy gain in
our ablations; additive conditioning alone is weaker.

\subsection{Alternating intra/inter-view attention}
Fusion alternates, for $T{=}2$ rounds, between two operators:
\emph{(i) intra-view} self-attention with AdaLN pose conditioning, which refines each
view's tokens independently under its own pose; and \emph{(ii) inter-view} attention
that, for each spatial token, attends across the $N$ views. Alternating the two lets
information move within and across orientations repeatedly:
\begin{equation}
\mathbf{F}^{(r+1)} = \mathrm{Inter}\!\big(\mathrm{Intra}(\mathbf{F}^{(r)};\{\mathbf{p}_i\})\big),
\qquad r=0,\dots,T{-}1 .
\end{equation}
A channel-stacked CNN, by contrast, must commit to a single fixed mixing of the
orientations and saturates as $N$ grows (\S\ref{sec:results}).

\subsection{Ray--microphone geometric cross-attention}
To read out an ERP depth map we lift a grid of output ray directions
$\{\hat{\mathbf{r}}_j\}_{j=1}^{R}$ (elevation $\times$ azimuth) and let each ray
cross-attend to the fused multi-view tokens. Crucially, the attention logits carry a
deterministic per-$(\text{ray},\text{view})$ geometric bias: for ray $\hat{\mathbf{r}}_j$
and view $i$ we rotate the ray into that microphone's frame,
$\hat{\mathbf{r}}_j^{(i)}=R(\psi_i)^{\!\top}\hat{\mathbf{r}}_j$, and score its agreement
with the ear axis $\hat{\mathbf{a}}_i=R(\psi_i)\,\hat{\mathbf{x}}$,
\begin{equation}
\mathrm{logit}_{j,i} = \frac{\mathbf{q}_j^{\!\top}\mathbf{k}_i}{\sqrt{C}}
      + b\big(\hat{\mathbf{r}}_j^{(i)},\,\epsilon_i\,\hat{\mathbf{a}}_i\big),
\end{equation}
so a ray preferentially draws on the views whose orientation actually illuminates
that direction. This is the mechanism that turns \emph{known pose} into a usable
inductive bias.

\subsection{Full-resolution multi-scale decoder}
The decoder upgrades the coarse $16\times32$ ray tokens to a native $256\times512$
depth map. A first stage upsamples the lift grid $ (L_h,L_w)\!\to\!(2L_h,2L_w)$ and
injects a \emph{fine} ray-lift skip (the ``multi-scale lift''), after which a stack of
ERP-aware residual blocks---2-D convolutions with \emph{circular} padding along the
azimuth so the panorama wraps seamlessly---progressively upsample to full resolution.
A final $1{\times}1$ head and sigmoid give radial depth
$\mathbf{d}=d_{\max}\cdot\sigma(\cdot)$. Predicting natively at $256\times512$ (rather
than upsampling a coarse map) is what makes the output directly comparable to the
ground truth without an interpolation penalty.

\subsection{Training objective}
The model is trained end-to-end with a single masked $L_1$ loss on valid pixels,
\begin{equation}
\mathcal{L}=\frac{\sum_{u} m_u\,\lvert \mathbf{d}_u-\mathbf{d}^{\star}_u\rvert}{\sum_u m_u},
\end{equation}
where $m$ is the valid-depth mask. We keep an exponential moving average
($0.999$) of the weights for evaluation. Auxiliary objectives (depth-bin
classification, aleatoric uncertainty, SILog, gradient/normal terms) were tested and
did not improve test accuracy, so the final model uses none. Because poses are known,
the same trained model generalises zero-shot to arbitrary orientation subsets by
simply passing the corresponding $\{\mathbf{p}_i\}$ at inference.
